---
title: "P8105 Fall 2023 Final Project Proposal"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(haven)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(reshape2)
library(knitr)
```

```{r include=FALSE}
import_df= function(path){
  read_xpt(path)
}
obesity=
  import_df(path='data/P_BMX.XPT') |> 
  full_join((import_df(path='data/P_DEMO.XPT')), by='SEQN') |>
  full_join((import_df(path='data/P_PAQ.XPT')), by='SEQN')|>
  full_join ((import_df(path='data/P_DBQ.XPT')), by='SEQN') |> 
  select (SEQN,RIAGENDR,RIDAGEYR,DMDMARTZ,INDFMPIR,RIDRETH3,DMDEDUC2,PAD680,BMXBMI,DBD900,DBD910) |> 
  filter(
    !(DMDMARTZ %in% c('77','99','.') ),
    !(INDFMPIR == '.'),
    !(DMDEDUC2 %in% c('7','9','.') ),
    !(PAD680 %in% c('7777', '9999', '.')),
    !(BMXBMI =='.'),
    !(DBD900 %in% c('7777', '9999', '.')),
    !(DBD910 %in% c('7777', '9999', '.'))
  ) |> 
  rename(
    gender=RIAGENDR,
    age=RIDAGEYR,
    marital_status=DMDMARTZ,
    income_to_poverty=INDFMPIR,
    race=RIDRETH3,
    education=DMDEDUC2,
    sedentary_activity=PAD680,
    bmi=BMXBMI,
    freq_fast_food=DBD900,
    freq_frozen=DBD910
  ) |> 
  mutate(
    gender = case_match(
      gender,
      1~"Male",
      2~"Female"),
    gender = as.factor(gender),
  marital_status= case_match(
    marital_status,
      1~"Married",
      2~"Widowed/Divorced/Separated",
      3~"Never married"
  ),
    race= case_match(
    race,
      1~"Mexican American",
      2~"Other Hispanic",
      3~"White",
      4~"Black",
      5~"Asian",
      6~"Other"
     ),
      education= case_match(
    education,
      1~"Less than 9th grade",
      2~"9-11th grade",
      3~"High school graduate" ,
      4~"Some college or AA degree",
      6~"College graduate or above"
     ),
   bmi =case_when(
    bmi<30 ~ 'normal',
    bmi>=30 ~'obese'
  )
  )
  
```


## Graphs

### Obesity Prevalence by Race and Gender
```{r}
obesity_race_filtered = obesity |>  
  filter(!is.na(race) & race != "NA")

ggplot(obesity_race_filtered, aes(x = race, fill = gender)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("pink", "lightblue")) +
  labs(title = "Obesity Prevalence by Race and Gender",
       x = "Race",
       y = "Count") +
  theme_minimal()
```

The bar chart displays the prevalence of obesity across different racial categories, differentiated by gender. It shows that the White population has the highest count of obesity cases, with a notable number of cases in both genders. The Black population follows, with females being higher than males, whereas the "Other Hispanic" categories have the lowest count of obesity cases.

### Age Distribution Among Obese Individuals
```{r}
ggplot(subset(obesity, bmi == "obese"), aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "pink", color = "lightblue") +
  labs(title = "Age Distribution Among Obese Individuals",
       x = "Age",
       y = "Count") +
  theme_minimal()
```

The histogram indicates that obesity is present across all age groups, with a relatively uniform distribution from young adulthood to late middle age. There's a significant peak in obesity cases among individuals in their late 50s to early 60s. After this peak, the number of obese individuals declines sharply for those in the older age groups.

### Income to Poverty Ratio Among Obese Individuals
```{r}
ggplot(subset(obesity, bmi == "obese"), aes(y = income_to_poverty, x = "", fill = bmi)) +
  geom_boxplot() +
  scale_fill_manual(values = c("pink")) +
  labs(title = "Income to Poverty Ratio Among Obese Individuals",
       x = "",
       y = "Income to Poverty Ratio") +
  theme_minimal()
```

The box plot shows the distribution of the income to poverty ratio among obese individuals. The data is spread across a range of ratios, mostly concentrated between 1 and 4, with the median around 2. This suggests that individuals with obesity come from a range of economic backgrounds, but there's a tendency towards lower income to poverty ratios.

### Histogram of Sedentary Activity
```{r}
obesity_sedentary_activity_filtered = obesity |> 
  filter(!is.na(sedentary_activity))

ggplot(obesity_sedentary_activity_filtered, aes(x = sedentary_activity)) +
  geom_histogram(binwidth = 1, fill = "pink", color = "lightblue") +
  labs(title = "Sedentary Activity Distribution Among Obese Individuals",
       x = "Sedentary Activity",
       y = "Count") +
  theme_minimal()
```

The majority of obese individuals report sedentary activity levels between 0 to around 500 minutes, with frequency decreasing as the amount of sedentary activity increases. There are notable peaks at certain intervals which may indicate common reporting increments (～250 minutes). Very few individuals report very high levels of sedentary activity, as seen by the low counts past 1000 minutes.

## Baseline characterics
```{r}
library(dplyr)
library(readxl)
```

```{r gender}
gender_summary <- obesity %>%
  count(gender) %>%
  mutate(Percentage = n / sum(n) * 100)

kable(gender_summary, format = "markdown", caption = "Gender Summary")
```

```{r marital_status}
marital_status_summary <- obesity %>%
  count(marital_status) %>%
  mutate(Percentage = n / sum(n) * 100)

kable(marital_status_summary, format = "markdown", caption = "Marital Status Summary")
```

```{r age}
age_summary <- obesity %>%
  summarise(
    Count = n(),
    Unique = n_distinct(age),
    Mean = mean(age, na.rm = TRUE),
    Std_Dev = sd(age, na.rm = TRUE),
    Min = min(age, na.rm = TRUE),
    `25th_Pctl` = quantile(age, 0.25, na.rm = TRUE),
    Median = median(age, na.rm = TRUE),
    `75th_Pctl` = quantile(age, 0.75, na.rm = TRUE),
    Max = max(age, na.rm = TRUE)
  )

kable(age_summary, format = "markdown", caption = "Age Summary")
```

```{r income}
income_to_poverty_summary <- obesity %>%
  summarise(
    Count = n(),
    Unique = n_distinct(income_to_poverty),
    Mean = mean(income_to_poverty, na.rm = TRUE),
    Std_Dev = sd(income_to_poverty, na.rm = TRUE),
    Min = min(income_to_poverty, na.rm = TRUE),
    `25th_Pctl` = quantile(income_to_poverty, 0.25, na.rm = TRUE),
    Median = median(income_to_poverty, na.rm = TRUE),
    `75th_Pctl` = quantile(income_to_poverty, 0.75, na.rm = TRUE),
    Max = max(income_to_poverty, na.rm = TRUE)
  )

kable(income_to_poverty_summary, format = "markdown", caption = "Income to Poverty Summary")
```

```{r education}
education_summary <- obesity %>%
  count(education) %>%
  mutate(Percentage = n / sum(n) * 100)

kable(education_summary, format = "markdown", caption = "Education Summary")
```


```{r race}
race_summary <- obesity %>%
  count(race) %>%
  mutate(Percentage = n / sum(n) * 100)

kable(race_summary, format = "markdown", caption = "Race Summary")
```

```{r sedentary_activity, echo=FALSE, message=FALSE}
sedentary_activity_summary <- obesity %>%
  summarise(
    Count = n(),
    Mean = mean(sedentary_activity, na.rm = TRUE),
    Std_Dev = sd(sedentary_activity, na.rm = TRUE),
    Min = min(sedentary_activity, na.rm = TRUE),
    `25th_Pctl` = quantile(sedentary_activity, 0.25, na.rm = TRUE),
    Median = median(sedentary_activity, na.rm = TRUE),
    `75th_Pctl` = quantile(sedentary_activity, 0.75, na.rm = TRUE),
    Max = max(sedentary_activity, na.rm = TRUE)
  )

kable(sedentary_activity_summary, format = "markdown", caption = "Sedentary Activity Summary")
```

```{r BMI}
bmi_summary <- obesity %>%
  count(bmi) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  ungroup() 

kable(bmi_summary, format = "markdown", caption = "BMI Summary")
```

```{r fast_food}

# Create age groups
obesity <- obesity %>%
  mutate(age_group = case_when(
    age >= 18 & age <= 25 ~ "18-25",
    age > 25 & age <= 35 ~ "26-35",
    age > 35 & age <= 45 ~ "36-45",
    age > 45 & age <= 55 ~ "46-55",
    age > 55 & age <= 65 ~ "56-65",
    age > 65 ~ "66+",
    TRUE ~ "Unknown"
  ))

fast_food_summary <- obesity %>%
  group_by(gender, age_group) %>%
  summarise(Avg_Fast_Food = mean(freq_fast_food, na.rm = TRUE))

kable(fast_food_summary, format = "markdown", caption = "Fast Food Frequency Summary by Age Group and Gender")

```

```{r frozen_food}
# Summarize by gender and age group
frozen_food_summary <- obesity %>%
  group_by(gender, age_group) %>%
  summarise(Avg_Frozen_Food = mean(freq_frozen, na.rm = TRUE)) 

kable(frozen_food_summary, format = "markdown", caption = "Frozen Food Frequency Summary by Age Group and Gender")
```


